---
title: "Homework 1 - Monte Carlo Methods"
author: "Jared Garfinkel - jsg2145"

output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1
The standard Laplace distribution has density $$f(x) = 0.5e^{-|x|}, x \in
(-\infty, \infty)$$. Please provide an algorithm that uses the inverse transformation method to generate a random sample from this distribution. Use the $U(0,1)$ random number generator in  $\em{\bf{R}}$, write a  $$\em{\bf{R}}$$-function to implement the algorithm. Use visualization tools to validate your algorithm (i.e., illustrate whether the random numbers generated from your function truely follows the standard Laplace distribution.)

# Answer: your answer starts here...

```{r }
#Your R codes/functions 
```

#Problem 2

Use the inverse transformation method to  derive an algorithm for generating a Pareto random number with $U\sim U(0,1)$, where the Pareto random number has a probability density function 
     $$f(x; \alpha, \gamma)=\frac{\gamma\alpha^{\gamma}}{x^{\gamma+1}} I\{x\ge \alpha\}$$ with two parameters $\alpha >0$ and $\gamma>0$. Use visualization tools to validate your algorithm (i.e., illustrate whether the random numbers generated from your function truely follows the target distribution.)


# Answer:  your answer starts here...
```{r }
#Your R codes/functions 
```

#Problem 3

Construct an algorithm for using the acceptance/rejection method to generate 100 pseudorandom
 variable from the pdf
 $$f(x) = \frac{2}{\pi \beta^2} \sqrt{\beta^2-x^2}, \,\, -\beta \le x \le \beta.$$
The simplest choice for $g(x)$ is the $U(-\beta, \beta)$ distribution but other choices are possible as well. Use visualization tools to validate your algorithm (i.e., illustrate whether the random numbers generated from your function truely follows the target distribution.)

# Answer:  your answer starts here...
```{r }
#Your R codes/functions 
```


#Problem 4

Develop two Monte Carlo methods for the estimation of $\theta=\int_0^1 e^{x^2} dx$ and implement in \em{\bf{R}} .

# Answer:  your answer starts here...
```{r }
#Your R codes/functions 
```

#Problem 5
Show that in estimating $\theta=E\sqrt{1-U^2}$ it is better to use $U^2$ rather than $U$ as the control variate, where $U\sim U(0,1)$.  To do this, use simulation to approximate the necessary covariances. In addition, implement your algorithms in  \em{\bf{R}}.

# Answer:  your answer starts here...
```{r }
#Your R codes/functions 
```

#Problem 6
Obtain a Monte Carlo estimate of
$$\int_1^\infty \frac{x^2}{\sqrt{2\pi}} e^{-\frac{x^2}{2}} dx$$
by importance sampling and evaluate its variance. Write a  \em{\bf{R}} function to implement your procedure.

# Answer:  your answer starts here...
```{r }
#Your R codes/functions 
```


